---
title: "Statistical Inference courser project "
author: "Jungsun Lee"
output: pdf_document
---

```{r setup, include=FALSE}
library(ggplot2)
library(dplyr)
```

# Part 1: Simulation Exercise Instructions
The project requires to investigate the exponential distribution in R and compare it with the Central Limit Theorem. The number of simulation, trial and lambda are given by the project instruction(simulation=1000,number=40,lambda=0.2). To get the same result,Here set the seed. 

```{r simulation,eco=FALSE}
set.seed(123)
n=40
lambda = 0.2
sim<-1000
expmean<-1/lambda
expsd<-1/lambda
mns = NULL
for (i in 1 : 1000) mns = c(mns, mean(rexp(n,lambda)))-> expsim
expsim<-as.data.frame(expsim)
```

## comparison of simulation and theoretical mean
plot the simulation result with simulation and theoretical means
```{r plot1}
ggplot(expsim,aes(x=expsim))+geom_histogram(aes(y=..density..),binwidth = lambda,color='black',fill='skyblue')+
  labs(x='means',title='Historgram of simulation ')+
  geom_vline(aes(xintercept=mean(expsim)),color="red")+
  geom_vline(aes(xintercept=expmean),color="green")

```

## Mean and variance 
The following figures are comparison of mean and variance of simulation result and theoretical calculus.
```{r comparison}
simulation_mean<-mean(expsim$expsim)
theoretical_mean<-expmean
simulation_var<-(sd(expsim$expsim))^2
theoratical_var<-(expsd/sqrt(n))^2
comparsion<-data.frame('simulation'=c(simulation_mean,simulation_var),
                       'theoratical'=c(theoretical_mean,theoratical_var),
                       row.names = c('mean','variance'))
comparsion
```

## Normality 
Whether the simulation distribution shows approximately normal or not, the figure below will show normal distribution with simulation result. 
```{r normality test}
ggplot(expsim,aes(x=expsim))+geom_histogram(aes(y=..density..),binwidth = lambda,color='black',fill='skyblue')+
  labs(x='means',title='Historgram of simulation ')+
  stat_function(fun = dnorm,args = list(mean=simulation_mean,sd=sqrt(simulation_var)),color='red',size=1.5)
```
As seen above, the simulation distribution is approximately normal distribution and follows the CLT

