---
title: "Reproducible project2"
author: "Jungsun Lee"
date: '2020 12 20 '
output: html_document
---
# Synopsis
This is a course project from Jonhs Hopkins's Data science specilazation. The questions of this assignment are following:
1.Across the United States, which types of events (as indicated in the EVTYPE variable) are most harmful with respect to population health?
2.Across the United States, which types of events have the greatest economic consequences?
The data is about damages from storms and severe weather in the US both economic and human lives cost. 
The questions are most bad consequences in health and economy so I will show top 10 types of events for each cases. 
```{r, eco=FALSE}
Sys.setlocale("LC_ALL", "English")
```
```{r }
library(ggplot2)
library(dplyr)
library(tidyr)
```

# Data loading 

```{r data downloading}
url<-'https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2FStormData.csv.bz2'
download.file(url,destfile ='repdata%2Fdata%2FStormData.csv.bz2')
stormdf<-read.csv('repdata%2Fdata%2FStormData.csv.bz2')
dim(stormdf)
```

# Data processing

## 1. Examine data name

```{r data name}
colnames(stormdf)
```

## 2. Data Subset
Here, subset the data for question 1 which asks most harmful types of events in population health.

```{r data subset 1}
# select the data
stormdf%>%select(EVTYPE,FATALITIES,INJURIES)->health 

# check the class type
str(health) 

 #change the class type to numeric
as.numeric(health$FATALITIES)->health$FATALITIES
as.numeric(health$INJURIES)->health$INJURIES 
```

Next, subset the data for question 2 which is types of events brought greatest damage to economy in US.

```{r data subset 2-1}
# select the data
stormdf%>%select(EVTYPE,PROPDMG,PROPDMGEXP,CROPDMG,CROPDMGEXP)->econ
as.numeric(econ$PROPDMG)->econ$PROPDMG
as.numeric(econ$CROPDMG)->econ$CROPDMG
str(econ)
```

Economic damages are divided into Property damages and Crop damages and those have damage amount and its exponents.It needs to convert exponent columns to actual exponents 

```{r data subset 2-2}
table(econ$PROPDMGEXP)
table(econ$CROPDMGEXP)

econ%>%mutate(prop=ifelse(PROPDMGEXP=="-",PROPDMG* 10^0,
                 ifelse(PROPDMGEXP=="?", PROPDMG*10^0,
                 ifelse(PROPDMGEXP=="+", PROPDMG*10^0,
                 ifelse(PROPDMGEXP=="0", PROPDMG*10^0,
                 ifelse(PROPDMGEXP=="1", PROPDMG*10^1,
                 ifelse(PROPDMGEXP=="2", PROPDMG*10^2,
                 ifelse(PROPDMGEXP=="3", PROPDMG*10^3,
                 ifelse(PROPDMGEXP=="4", PROPDMG*10^4,
                 ifelse(PROPDMGEXP=="5", PROPDMG*10^3,
                 ifelse(PROPDMGEXP=="6", PROPDMG*10^6,
                 ifelse(PROPDMGEXP=="7", PROPDMG*10^7,
                 ifelse(PROPDMGEXP=="8", PROPDMG*10^8,
                 ifelse(PROPDMGEXP=="H", PROPDMG*10^2,
                 ifelse(PROPDMGEXP=="h", PROPDMG*10^2,
                 ifelse(PROPDMGEXP=="K", PROPDMG*10^3,
                 ifelse(PROPDMGEXP=="M", PROPDMG*10^6,
                 ifelse(PROPDMGEXP=="m", PROPDMG*10^6,
                 ifelse(PROPDMGEXP=="B", PROPDMG*10^9,0)))))))))))))))))))->econ

              
econ%>%mutate(crop=ifelse(CROPDMGEXP=="",CROPDMG*10^0,
          ifelse(CROPDMGEXP=="?",CROPDMG*10^0,
          ifelse(CROPDMGEXP=="0", CROPDMG*10^0,
          ifelse(CROPDMGEXP=="2", CROPDMG*10^2,
          ifelse(CROPDMGEXP=="k", CROPDMG*10^3,
          ifelse(CROPDMGEXP=="K", CROPDMG*10^3,
          ifelse(CROPDMGEXP=="M", CROPDMG*10^6,
          ifelse(CROPDMGEXP=="m", CROPDMG*10^6,
          ifelse(CROPDMGEXP=="B", CROPDMG*10^9,0))))))))))->econ

head(econ)

```


## 3. Calculating columns

calculating total numbers of fatality and injuries of each type of events 
```{r h}
health%>%group_by(EVTYPE)%>%summarise(fatality=sum(FATALITIES,na.rm = TRUE),injuries=sum(INJURIES,na.rm = TRUE))%>%
  arrange(desc(fatality,injuries))->h
head(h)
```

calculating total amounts of damages of properties and crops by events
```{r e}
econ%>%select(EVTYPE,prop,crop)%>%group_by(EVTYPE)%>%
  summarise(propdamage=sum(prop,na.rm = TRUE),cropdamage=sum(crop,na.rm = TRUE))%>%
  arrange(desc(propdamage,cropdamage))->econdamage
head(econdamage)
```

# Result

## 1.The 10 most harmful to population health
making tidy data by selecting 10 most harmful events
```{r tidyh}
h%>%filter(row_number()<=10)%>%gather(key="harmful_health",value='total number',-EVTYPE)->tidyh
```

```{r plot1}
ggplot(tidyh)+geom_bar(stat = 'identity',aes(x=reorder(EVTYPE,-`total number`),y=`total number`,fill=harmful_health ), position="dodge")+
  theme(axis.text.x = element_text(angle=45, hjust=1))+labs(x='Types of Events',y='Total number')+
  ggtitle('The top 10 most harmful types of event in health ')
```

## 2. 10 Greatest econimic damages
making tidy data
```{r tidye}
econdamage%>%filter(row_number()<=10)%>%gather(key="DamageInEcon",value='Total Amount',-EVTYPE)->tidye
```

```{r plot2}
ggplot(tidye)+geom_bar(stat ='identity',aes(x=reorder(EVTYPE,-`Total Amount`),y=`Total Amount`,fill=DamageInEcon ), position="dodge")+
                         theme(axis.text.x = element_text(angle=45, hjust=1))+labs(x='Types of Events',y='Total Amount')+
                         ggtitle('The top 10 most Damaged types of event in Economy')
```