---
title: "Intro to Prob and Data with R test"
output: html_document
---
part1. Data

*discuss whether random sampling was used
Yes, It collects the data in randomly sampling 
From the BRFSS overview, interviewers collect data from a randomly selected adults in a household in landline telephone survey and people participating cellular telephone surveys are those who living in a private residence or college housing. Therefore this survey collect the data in randomly sampling.

* discuss whether random assignment was used
No, this is not a random assignment
collected data is answers from questionnaires of people who participating in the surveys, which mean it is observational not assigned.

-> Hence, This is not causal but generalizable
part2 reserach questions

1.Is there any relationship between sleeping hours and poor physical or mental health?
2.how much percetage of people take part in a depress disorder in each sleeping group?
among those who ever told in a depression or not in each sleep group, how many days do they feel poor physical or mental health?
3.among those who ever told in a depression or not in each sleep group,how many drinks did you drink on the average?
```{r load_libraries, message=FALSE}
library(dplyr)
library(ggplot2)
library(statsr)
```
```{r set the directory}
#getwd()
#setwd("C:/Users/Jungsun/Desktop/coursera/statistics with r/1. Introduction pro and data with r")
#getwd()
```
# read the data
```{r load the_data}
data<-load("brfss2013_test.rdata")
data<-brfss2013
```
# basic understanding of data
```{r get_the_data}
#how many times people usually sleep?
#remove NA values and 
df<-data%>%filter(!is.na(sleptim1)&sleptim1<=24)#remove outlier
sl<-df%>%group_by(sleptim1)%>%summarise(count=n())
df%>%summarise(mean_sl=mean(sleptim1),sd_sl=sd(sleptim1))
```

```{r plot for sleeping hours on average}
ggplot(data=sl,aes(x=factor(sleptim1),y=count))+geom_point()
```

As we can see from the sleeping hour data, it has unimodal and right skew.

Reserach question 1.Is there any relationship between sleeping hours and poor physical or mental health?
```{r boxplot of each group }
df2<-df%>%filter(!is.na(poorhlth)&poorhlth<30) #remove NA values and over 30 
df2%>%summarise(mean_p=mean(poorhlth),sd_p=sd(poorhlth))
df_sub<-subset(df2,select = c(sleptim1,poorhlth,addepev2))
df_sub%>%group_by(sleptim1)%>%summarise(mean_slp=mean(poorhlth),sd_slp=sd(poorhlth))
```

#plot the data
```{r plot for poor health of each sleeping}
ggplot(data=df_sub,aes(x=factor(sleptim1),y=poorhlth))+geom_boxplot()
```
From the questionnaire of the survey, it asked  Poor Physical Or Mental Health During the past 30 day. Therefore, remove poor physical and mental health over 30. 
Each sleeping hour, calculate the average of poor health.We cannot make any causal relationship between sleeping hours and health condition. but we can check how many times on average people feel poor health under certain sleeping hours.
As we can see from the plot, people who sleep about 5 to 9 hours have relatively low average and standard deviation. 
It means people who sleep about 5 to 9 hours feel less physical or mental weakness compared to those who sleep less or more.

Reserach question 2.1.how much percentage of people take part in a depress disorder in each sleeping group?

This question is whether people who have ever been suffered from depression disorder or not have different their sleeping hours on average.
```{r ever having depression disorder  }
df_sub<-df_sub%>%filter(!is.na(addepev2))
df_sub%>%group_by(sleptim1)%>%summarise(dep_rate=sum(addepev2=="Yes")/n())
```

#plot the data
```{r plot for depression disorder}
ggplot(data=df_sub,aes(x=factor(sleptim1),fill=addepev2))+geom_bar()
```

The depression rate among those who have ever been in a depression disorder decreases until 7 hours per day and increase after that, Which is shown similar result above research question.

Research question 2.2.Among those who ever told in a depression or not in each sleep group, how many days do they feel poor physical or mental health?
# poor health condition days on average those who have suffered from depression disorder by each sleeping hour
```{r comparision of poor health conditon between those who have suffered from depression disorder by each sleeping hour}
dep<-df_sub%>%filter(!is.na(addepev2)&addepev2=="Yes")
dep_sm<-dep%>%group_by(sleptim1)%>%summarise(mean_p=mean(poorhlth))  
dep_sm
```
# plot the data
```{r plot for comparision of poor health }
ggplot(data=dep,aes(x=factor(sleptim1),y=poorhlth))+geom_boxplot()
```
# poor health condtion days on average who never have depression diseases
```{r comparision of poor health conditon between those who never have depression disorder by each sleeping hour}
no_dep<-df_sub%>%filter(!is.na(addepev2)&addepev2=="No")
nodep_sm<-no_dep%>%group_by(sleptim1)%>%summarise(mean_dp=mean(poorhlth)) 
nodep_sm
```
# plot the data
```{r plot for poor health condition not depress}
ggplot(data=no_dep,aes(x=factor(sleptim1),y=poorhlth))+geom_boxplot()
```
# To see the comparision in one graph ( merge the data)
```{r comparision of poor health conditon}
dep_sub<-merge(dep_sm,nodep_sm,all = TRUE)
dep_sub<-na.omit(dep_sub)
g<-ggplot(data=dep_sub,aes(x=factor(sleptim1)))+geom_point(aes(y=mean_p),color="darkblue")
g<-g+geom_point(aes(y=mean_dp),color="darkred")
g
```
The graph shows us that people who have ever been in a depression disorder have more poor physical or mental condition on average in each sleep hour. It is not clear people still suffer from depression disorder or not among those who have ever told in a depression problem. But we can see clear relation between poor health condition and depression disorder.

Research question 3.Does the depression problem show average alcohol consumption less or higher in each sleeping hour?
This cast a question about people who have ever been in depression disorder might suffer other bad habit or less control of them. 
To see this, it compares their consumption of alcohol on average in every sleep hour
The reason grouping by every sleeping hour is that it assumes people might less sleep due to drink alcohol or consumption of alcohol affects sleeping hours.
#select the data 
```{r select the data and filter by depression }
df_sub2<-subset(data,select=c(sleptim1,addepev2,avedrnk2))
df_sub2<-na.omit(df_sub2) # remove all NA's 
dep2<-df_sub2%>%filter(addepev2=="Yes")%>%group_by(sleptim1)%>%summarise(mean_dr=mean(avedrnk2))
dep2
```

```{r no depress}
no_dep2<-df_sub2%>%filter(addepev2=="No")%>%group_by(sleptim1)%>%summarise(mean_nodr=mean(avedrnk2))
no_dep2
```
# plot the data
```{r comparison of comsumption of al}
g2<-ggplot()+geom_point(data=dep2,aes(x=factor(sleptim1),y=mean_dr))#depression
g2<-g2+geom_point(data=no_dep2,aes(x=factor(sleptim1),y=mean_nodr,color="darkred"))# no depression
g2
```
This graph does not show clear tendency between depression and average drinking alcohol in every sleeping hour.
Therefore, we can assume that depression problem might not relate to average consumption of alcohol But it only compares average consumption rate. It might needs more research between them. 
